---
title:
  "Basic Walkthrough"
description: >
  This vignette describes how to train a Falcata model for binary classification.
output:
  markdown::html_format:
    options:
      toc: true
      number_sections: true
vignette: >
  %\VignetteIndexEntry{Basic Walkthrough}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE
  , comment = "#>"
  , warning = FALSE
  , message = FALSE
)
```

## Introduction

Welcome to the world of [Falcata](https://github.com/MechaFauna-ai/Falcata), a highly efficient gradient boosting implementation (Ke et al. 2017).

```{r}
library(falcata)
```

```{r, include=FALSE}
# limit number of threads used, to be respectful of CRAN's resources when it checks this vignette
data.table::setDTthreads(1L)
setFalcataThreads(2L)
```

This vignette will guide you through its basic usage. It will show how to build a simple binary classification model based on a subset of the `bank` dataset (Moro, Cortez, and Rita 2014). You will use the two input features "age" and "balance" to predict whether a client has subscribed a term deposit.

## The dataset

The dataset looks as follows.

```{r}
data(bank, package = "falcata")

bank[1L:5L, c("y", "age", "balance")]

# Distribution of the response
table(bank$y)
```

## Training the model

The R-package of Falcata offers two functions to train a model:

- `lgb.train()`: This is the main training logic. It offers full flexibility but requires a `Dataset` object created by the `lgb.Dataset()` function.
- `falcata()`: Simpler, but less flexible. Data can be passed without having to bother with `lgb.Dataset()`.

### Using the `falcata()` function

In a first step, you need to convert data to numeric. Afterwards, you are ready to fit the model by the `falcata()` function.

```{r}
# Numeric response and feature matrix
y <- as.numeric(bank$y == "yes")
X <- data.matrix(bank[, c("age", "balance")])

# Train
fit <- falcata(
  data = X
  , label = y
  , params = list(
    num_leaves = 4L
    , learning_rate = 1.0
    , objective = "binary"
  )
  , nrounds = 10L
  , verbose = -1L
)

# Result
summary(predict(fit, X))
```

It seems to have worked! And the predictions are indeed probabilities between 0 and 1.

### Using the `lgb.train()` function

Alternatively, you can go for the more flexible interface `lgb.train()`. Here, as an additional step, you need to prepare `y` and `X` by the data API `lgb.Dataset()` of Falcata. Parameters are passed to `lgb.train()` as a named list.

```{r}
# Data interface
dtrain <- lgb.Dataset(X, label = y)

# Parameters
params <- list(
  objective = "binary"
  , num_leaves = 4L
  , learning_rate = 1.0
)

# Train
fit <- lgb.train(
  params
  , data = dtrain
  , nrounds = 10L
  , verbose = -1L
)
```

Try it out! If stuck, visit Falcata's [documentation](https://github.com/MechaFauna-ai/Falcata/tree/master/R-package) for more details.

```{r, echo = FALSE, results = "hide"}
# Cleanup
if (file.exists("falcata.model")) {
  file.remove("falcata.model")
}
```

## References

Ke, Guolin, Qi Meng, Thomas Finley, Taifeng Wang, Wei Chen, Weidong Ma, Qiwei Ye, and Tie-Yan Liu. 2017. "LightGBM: A Highly Efficient Gradient Boosting Decision Tree." In Advances in Neural Information Processing Systems 30 (NIPS 2017).

Moro, Sérgio, Paulo Cortez, and Paulo Rita. 2014. "A Data-Driven Approach to Predict the Success of Bank Telemarketing." Decision Support Systems 62: 22–31.
